---
title: "Random week sample generator"
author: "Vivien Benert"
date: "2023-02-06"
output:
  word_document: default
  pdf_document: default
  html_document: default
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
# library(dplyr)
# library(lubridate)
```

# Function to draw artificial random week from sample period

The function needs a start date and an end date, provided in the format yyyy-mm-dd. Additionally, there is the option to specify whether the sample should be drawn from the whole week or only from weekdays. Default is set to take the whole week as the basis for the random sample, including weekends. If weekends should be excluded (i.e. if the random sample should only inlcude mondays to fridays, set the optional weekends-argument to FALSE).

```{r}
sample_week <- function(start.date, end.date, weekend=TRUE) {
 require(dplyr)
 require(lubridate)
  # Create sequence of days for sample period
  df <- as.data.frame(seq(ymd(start.date), ymd(end.date), by='days'))
  colnames(df) <- "date"
  
  # Create weekday and year variable with lubridate 
  df$weekday <- wday(df$date)
  df$year <- year(df$date)
  df$week <- week(df$date)
  
  # Select only weekdays, only if weekends=FALSE
  if (weekend==FALSE) {
    x <- 5
  } else {
    x <- 7
  }

  df <- df[which(df$weekday<=x),]
  
  # Subset years, returns a list of dfs
  df_list <- split(df, f=df$year)
  
  # Create empty data frame
  sample <- data.frame(matrix(ncol=4, nrow=0))
  colnames(sample) <- c("date", "weekday", "year", "week")

  for (i in 1:length(df_list)) {
    x <- df_list[[i]]
    z <- x %>% 
      group_by(week) %>% 
      sample_n(1) %>% 
      ungroup()
    sample <- rbind(sample, z)
    }

  return(sample)
}
```


```{r}
# Use function by providing start date and end date as yyyy-mm-dd format
# Don't forget to set.seed if sample needs to be reproducible.

# set.seed(12345)

# sample_week <- sample_week('1955-06-08', '2023-01-01')
# sample_weekdays <- sample_week('1989-05-12', '2022-10-30', weekend = FALSE)
```
